Design, Implementation and Characterization of a Sigma-Delta Analog-to-Digital Converter in FPGA

### 2016-04-10 extras/paper_en_ingles/Paper_ADC_Sigma-Delta_en_FPGA.tex
Comenzando traducción paper

%% TILDES
\usepackage[utf8]{inputenc}
%\usepackage[spanish]{babel}

@@

\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  \graphicspath{{./eps/} }%{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.eps} %.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  \graphicspath{{./eps/}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

@@

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{ADC $\Sigma\Delta$ en FPGA}

@@

\begin{abstract}
%\boldmath
En el desarrollo de esta publicación, se presenta un conversor analógico digital en una \textit{FPGA}, implementado bajo una topología Sigma-Delta la cual permite una realización con una cantidad muy pequeña de hardware externo. Se utiliza el buffer diferencial \textit{LVDS} como entrada y se emplean topologías de filtrado digital que aprovechan los recursos propios de la \textit{FPGA} buscando así obtener un conversor de bajo costo y consumiendo pocos recursos. Se desarrollan simulaciones y metodologías de medición para caracterizarlo y poder realizar la comparación con los resultados ideales. Como resultado del trabajo se ha obtenido un conversor analógico digital de 9 bits efectivos en bajas frecuencias, con una frecuencia de muestreo de 22KHz.

%En el desarrollo de esta publicación, se presenta un conversor analógico digital en una \textit{FPGA}, implementado bajo una topología Sigma-Delta la cual permite una realización con una cantidad muy pequeña de hardware externo. También, se buscaron distintas topologías de filtrado digital para aprovechar los recursos propios de la \textit{FPGA}. Finalmente, se desarrollaron simulaciones y metodologías de medición para poder caracterizarlo y poder realizar la comparación con los resultados ideales. \\

@@

\section{Introducci\'on}
% no \IEEEPARstart
Un conversor analógico digital sigma delta est\'a basado en los conceptos de \textit{Oversampling} y \textit{Noise Shaping} (distribuci\'on del ruido). El \textit{Oversampling} consiste en el muestreo de una señal a una frecuencia mayor que la frecuencia de Nyquist, logrando distribuir la potencia de ruido de cuantización (considerada constante) en un ancho de banda mayor. El \textit{Noise Shaping} se basa en la redistribución del ruido de cuantizaci\'on de forma tal que quede fuera de la banda de inter\'es, pudiendo ser disminuido como en el caso anterior (ver Fig.~\ref{fig:NS}). Estos dos efectos permiten eliminar la mayor parte del ruido de cuantización con un filtro pasabajos.
% El \textit{Oversampling} consiste en el muestreo de una señal a una frecuencia mayor que la frecuencia de Nyquist que, al considerar que la potencia de ruido de cuantización es constante y no depende de la frecuencia de muestreo, el ruido de cuantizaci\'on se distribuye en un ancho de banda mayor, permitiendo que este pueda ser posteriormente disminuido por medio de un filtro digital. El \textit{Noise Shaping} consiste en la redistribución del ruido de cuantizaci\'on de forma tal que quede fuera de la banda de inter\'es, pudiendo ser disminuido como en el anterior caso (ver Fig.~\ref{fig:NS}).

\begin{figure}[!b]
\centering
\includegraphics[width=1.7in]{Noise_Shaping.eps}
\caption{Arriba Oversampling. Abajo Oversampling+Noise Shapping}
\label{fig:NS}
\end{figure}

La topolog\'ia de un ADC sigma delta está representada en el diagrama en bloques de la Fig.~\ref{fig:SDTD}. La adición de ruido $N(z)$ es la modelización del ruido de cuantización. Analizando el diagrama, se puede obtener que:

\begin{align}
H_x(z)=\frac{H(z)}{1+H(z)} \qquad
H_n(z)=\frac{1}{1+H(z)}
\end{align}

donde $H_x$ y $H_n$ son las transferencias para la señal y para el ruido respectivamente. Siendo $H(z)$ un filtro pasabajos, $H_x(z)$ también será un filtro pasabajos (mientras sea $H(z)$ en la banda de paso mayor a 1) y $H_n(z)$ un filtro pasa altos. De esta forma se obtiene el \textit{Noise Shaping}.

\begin{figure}[!t]
\centering
\includegraphics[height=1in]{Sigma-Delta_Tiempo_Discreto.eps}
\caption{Diagrana en bloques del modulador $\Sigma\Delta$}
\label{fig:SDTD}
\end{figure}

Para el desarrollo del proyecto, se utilizó el kit de desarrollo Mojo v3 el cual posee una \textit{FPGA Xilinx Spartan6}. En la Fig.~\ref{fig:DBADC} se resume el sistema completo, donde el filtro CIC es un filtro de decimación que utiliza pocos recursos de la FPGA. Este realiza una decimación de 512 veces, mientras que el filtro FIR decima 4, dando un factor de decimación total de 2048 veces.
%\hfill mds
%\hfill January 11, 2007

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{Diagrama_Sistema_ADC.eps}
\caption{Diagrama en bloques del ADC}
\label{fig:DBADC}
\end{figure}

\subsection{Modulador $\Sigma\Delta$ en tiempo discreto}
Basado en el diagrama en bloques de la figura \ref{fig:SDTD}, un modulador de primer orden puede modelizarse con la siguiente transferencia:

\begin{align}
H(z)=\frac{z^{-1}}{1-z^{-1}}
\end{align}
De esto se obtiene que:
\begin{align}
H_x(z)=z^{-1} \qquad
H_n(z)=1-z^{-1}
\end{align}

Por lo tanto, la transferencia para la señal es simplemente un retardo, y la del ruido consta de la diferencia entre la muestra actual y la anterior.
Para implementar un filtro de orden k se suele recurrir a que su transferencia esté dada por:

\begin{align}
H_n(z)=\left( 1-z^{-1}\right)^{k}
\end{align}

\subsection{Ruido de cuantizaci\'on}
Suponiendo el ruido de cuantización como ruido blanco uniformemente distribuido entre -$\Delta$/2 y $\Delta$/2, siendo $\Delta$ el alto del escal\'on de cuantización se tiene que \cite{Tesis:Hellman}:

\begin{align}
S_q(f) &= \frac{\Delta^{2}}{12f_s}\\
P_{q}(f) &=\int\limits_{-f_s/2}^{f_s/2}S_q(f) \, |H_n(f) \times H_D(f)|^{2} df
\end{align}

Donde $S_q,P_q$ y $H_D$ son la densidad espectral de potencia de ruido, la potencia de ruido de cuantizaci\'on a la salida, y la transferencia del filtro digital respectivamente.
Teniendo en cuenta un filtro digital ideal con corte en $f_p$, y un modulador de orden k se obtiene que la relaci\'on se\~nal a ruido de cuantizaci\'on en funci\'on del \textit{oversampling ratio}
\footnote[1]
{
Se lo obtiene teniendo en cuenta que en el integrando para $OSR\gg 1$ es $f\leq f_p\ll f_s$ y $|H_n(e^{j2 \pi \frac{f}{f_s}})|^{2}= \sin(\pi \frac{f}{f_s})^{2k} \approx (\pi\frac{f}{f_s})^{2k}$ y que la potencia un seno de plena escala es $P_{sin}=\frac{1}{2} [(2^N-1)\frac{\Delta}{2}]^{2}$.
Usualmente aparece la formula con $2^{N}-1\approx2^N$.
} es:

@@

En la Fig.~\ref{fig:SQNR} se ilustra el resultado.

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{SQNR_ideal_segun_OSR}
%\def\svgwidth{2.5in}
%\input{Noise_Shaping2.eps}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
\caption{SQNR en funci\'on de OSR para moduladores de diferente orden}
\label{fig:SQNR}
\end{figure}
%\subsubsection{Subsubsection Heading Here}
%Subsubsection text here.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals/conferences use top floats
% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
% command of the stfloats package.

\section{Modulador $\Sigma\Delta$ en tiempo continuo}

A diferencia del modulador $\Sigma\Delta$ en tiempo discreto, el muestreo se realiza dentro del modulador, luego de la cuantización. Dicho cambio se aprecia en la Fig.~\ref{fig:SDTC} de la cual que se puede obtener la siguiente transferencia para la señal y para el ruido:

%En la Fig.~\ref{fig:SDTC} donde se reemplazó el cuantizador por su modelo de ruido aditivo, las transferencias para la se\~nal y para el ruido quedan:

\begin{align}
H_x(e^{sT_s}) &= \frac{H(s)}{1+H_{eq}(e^{sT_s})} \\
H_n(z) &= \frac{1}{1+H_{eq}(z)} \\
h_{eq}[n] &= (h\ast h_{DAC})(nT_s)
\end{align}

Donde la respuesta de lazo equivalente en tiempo discreto $H_{eq}(z)$ se la obtiene por la transformaci\'on de invarianza al impulso\cite{DSP:Pro-Man} de $H(s)H_{DAC}(s)$ dada por (11). Para un retenedor de orden cero queda como:

@@

Partiendo del modulador de primer orden en tiempo discreto y con la anterior transformaci\'on se obtiene:

\begin{align}
 H(s)&=\frac{1}{sT_s}\\
 H_x(e^{sT_s})&= \frac{1-e^{-sT_s}}{s}
\end{align}

Considerando que el filtro $H(s)$ es un integrador, la trasferencia para la se\~nal es un pasabajos con ceros en múltiplos de $f_s$, lo cual reduce notablemente el alias en la banda de paso.

\begin{figure}[!t]
\centering
\includegraphics[height=1in]{Sigma-Delta_Tiempo_Continuo}
\caption{Diag. en bloques del modulador $\Sigma\Delta$ en tiempo continuo}
\label{fig:SDTC}
\end{figure}

>>>>

%% TILDES
%\usepackage[utf8]{inputenc}
%\usepackage[spanish]{babel}

@@

\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  \graphicspath{{../paper/eps/} }%{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.eps} %.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  \graphicspath{{./eps/}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

@@

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Design, Implementation and Characterization of a Sigma-Delta Analog-to-Digital Converter in FPGA}

@@

\begin{abstract}
%\boldmath
While analog to digital converters are often part of microcontrollers, this doesn't happen with FPGA, forcing a developer to obtain an external converter for his projects. As a FPGA is a widely used tool for DSP, we developed a Sigma-Delta ADC in the FPGA, using very low external hardware. In this publication we describe the design, implementation and measure of this converter.

@@

\section{Introduction}
% no \IEEEPARstart

Sigma-delta ($\Sigma\Delta$) analog to digital converter is based on two main conceps: \textit{Oversampling} and \textit{Noise Shaping}.
\textit{Oversampling} consists on sampling a signal with a frequency higher than Nyquist's. This distributes the quantization noise in a bigger bandwidth.
\textit{Noise shaping} consists on redistributing the quantization noise in such a way that most of it will be out of the band of interest.
Both concepts are shown in Fig.~\ref{fig:NS}.
The modulator is followed by a digital low pass decimation filter, eliminating most of the in-band quantization noise. This allows converters with high ENOB.

\begin{figure}[!b]
\centering
\includegraphics[width=1.7in]{Noise_Shaping.eps}
\caption{Up: Oversampling. Low: Oversampling+Noise Shapping}
\label{fig:NS}
\end{figure}

The $\Sigma\Delta$ ADC topology is in the block diagram of Fig. ~\ref{fig:SDTD}. The quantization noise is modeled as additive white noise ($N(z)$). The transfers functions for the signal (STF) and the noise (NTF) are:

\begin{align}
STF(z)=\frac{H(z)}{1+H(z)} \qquad
NTF(z)=\frac{1}{1+H(z)}
\end{align}

Being H a low pass filter, STF will be a low pass filter, and NTF will be a high pass filter, obtaining the previously mentioned Noise Shaping.

\begin{figure}[!t]
\centering
\includegraphics[height=1in]{Sigma-Delta_Tiempo_Discreto.eps}
\caption{Block diagram of the $\Sigma\Delta$}
\label{fig:SDTD}
\end{figure}

% % Poner en materiales y metodos
For the develop of the project, we used the developing kit \textit{Mojo v3}, that incorpores a \textit{Xilinx Spartan 6 FPGA}.
% %

In ~\ref{fig:DBADC} we show the basic blocks of the converter, where we can distinguish the modulator (analog stage), and two digital filters (CIC and FIR). The CIC (Cascaded integrator–comb) is a decimation filter that can be implemented with few%low
FPGA resources. The decimation factor we used is 512. The FIR is the one that will eliminate the noise accumulated in the high frequencies. After the FIR, we decimate again with a factor of 4, obtaining a total decimation factor of 2048.

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{Diagrama_Sistema_ADC.eps}
\caption{Block diagram of the ADC}
\label{fig:DBADC}
\end{figure}

\subsection{Discrete-Time $\Sigma\Delta$ Modulator}

If H in Fig. 2 as an accumulator:

\begin{align}
H(z)=\frac{z^{-1}}{1-z^{-1}}
\end{align}

Using Eq. 1, we obtain the STF and NTF as:

\begin{align}
STF(z)=z^{-1} \qquad
NTF(z)=1-z^{-1}
\end{align}

Then the STF is just a delay, while the NTF is a high pass filter as is wanted.
To implement a k-order filter, the following NTF is used:

\begin{align}
NTF(z)=\left(1-z^{-1}\right)^{k}
\end{align}

\subsection{Quantization noise}

Assuming the quantization noise (QN) is white and uniformly distributed between -$\Delta/2$ and $\Delta/2$, being $\Delta$ the quantization step, it can be shown that:

% % poner una referencia

\begin{align}
S_q(f) &= \frac{\Delta^{2}}{12f_s}\\
P_{q}(f) &=\int\limits_{-f_s/2}^{f_s/2}S_q(f) \, |NTF(f) \times H_D(f)|^{2} df
\end{align}

where $S_q$ is the QN power spectral density and $P_q$ is the QN power after Noise Shapping and Digital Filtering, being $H_D$ the transfer function of the digital filter.
For an ideal digital filter with cutoff frequency FP, and a modulator of order k, the signal to quantization noise relationship is:

@@

where OSR is the oversampling ratio.

% %CAMBIAR FORMULA DE ARRIBA POR LA FORMULA APROXIMADA QUE APARECE SIEMPRE CON UNA REFERENCIA Y SIN ESTA EXPLICACION
%\footnote[1]
%{
%Se lo obtiene teniendo en cuenta que en el integrando para $OSR\gg 1$ es $f\leq f_p\ll f_s$ y $|H_n(e^{j2 \pi \frac{f}{f_s}})|^{2}= \sin(\pi \frac{f}{f_s})^{2k} \approx (\pi\frac{f}{f_s})^{2k}$ y que la potencia un seno de plena escala es $P_{sin}=\frac{1}{2} [(2^N-1)\frac{\Delta}{2}]^{2}$.
%Usualmente aparece la formula con $2^{N}-1\approx2^N$.
%} es:

The resulting SQNR is plotted in Fig.~\ref{fig:SQNR} as a function of the OSR for different filters orders.

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{SQNR_ideal_segun_OSR}
%\def\svgwidth{2.5in}
%\input{Noise_Shaping2.eps}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
\caption{SQNR as a function of de OSR for different filters orders}
\label{fig:SQNR}
\end{figure}

\section{Continuous-Time $\Sigma\Delta$ Modulator}

Unlike the discrete-time $\Sigma\Delta$ modulator, here the sampling is made inside the modulator, right after quantization (Fig.~\ref{fig:SDTC}). The STF and NTF are obtained as:

\begin{align}
STF(e^{sT_s}) &= \frac{H(s)}{1+H_{eq}(e^{sT_s})} \\
NTF(z) &= \frac{1}{1+H_{eq}(z)} \\
h_{eq}[n] &= (h\ast h_{DAC})(nT_s)
\end{align}

where the equivalent discrete-time $H_eq$ response is obtained with the impulse-invariance transform\cite{DSP:Pro-Man} of $H(s)H_{DAC}(s)$ defined in (11).
Using the DAC as a Zero-order hold is:

@@

With the first order discrete time modulator and the previous transformation, we obtain:

\begin{align}
 H(s)&=\frac{1}{sT_s}\\
 STF(e^{sT_s})&= \frac{1-e^{-sT_s}}{s}
\end{align}

Being H an integrator, and the STF a low pass filter with zeros in multiples of $f_s$, which reduces the alias in the passband.

\begin{figure}[!t]
\centering
\includegraphics[height=1in]{Sigma-Delta_Tiempo_Continuo}
\caption{Block diagram of a continuous-time $\Sigma\Delta$ modulator}
\label{fig:SDTC}
\end{figure}

% %Seguir desde aca
%A. Implementation
% %The simplest implementation for the modulator is an RC filter (fig. 6). The quantization is done by a LVDS input of the FPGA [6], commonly used for differnetial communications, and the time discretization is done by a flip flop.
